### 2016-05-04 p03-ch01-2015-monitoring/paper-esocc.tex
Add publication information to chapter abstracts
Fix ref/label for Chapter 04

\section*{Abstract}
\label{abstract}
In this paper, we introduce a  formal model of the  availability, budget compliance
and  sustainability of distributed services, where service   sustainability is
a new concept which arises as 
% an extension of 
the composition of 
service  availability and budget compliance.
The model formalizes a distributed  platform for monitoring the above
service characteristics in terms of a parallel composition of task automata,
where dynamically generated tasks model asynchronous events with deadlines.
The main result of this paper is a formal model to optimize and reason about service characteristics
through monitoring.
In particular, we use schedulability analysis of the
underlying timed automata to optimize and guarantee service sustainability.

@@

>>>>

\section*{Abstract}
\label{ch04:abstract}
In this paper, we introduce a  formal model of the  availability, budget compliance
and  sustainability of distributed services, where service   sustainability is
a new concept which arises as 
% an extension of 
the composition of 
service  availability and budget compliance.
The model formalizes a distributed  platform for monitoring the above
service characteristics in terms of a parallel composition of task automata,
where dynamically generated tasks model asynchronous events with deadlines.
The main result of this paper is a formal model to optimize and reason about service characteristics
through monitoring.
In particular, we use schedulability analysis of the
underlying timed automata to optimize and guarantee service sustainability.

@@

%

\subparagraph*{Conference Publication}
\emph{Lecture Notes in Computer Science, Volume 9306, Service Oriented Cloud Computing, 4th European Conference -- ESOCC 2015, Pages 125--140, DOI 
\smalltt{10.1007/978-3-319-24072-5\_9}}

%

### 2016-05-04 p03-ch01-2015-monitoring/sec02-related.tex
Fix ref/label for Chapter 04

\section{Related Work}
\label{sec:relatedwork}

@@

Hogben and Pannetrat examine in \cite{hogben2013defavail} the challenges of defining and measuring availability to support real-world service comparison and dispute resolution through SLAs.
They show how two examples of real-world SLAs would lead one service provider to report 0\% availability while another would report 100\% for the same system state history but using a different period of time.
The transparency that the authors attempt to reach is addressed in our work by the concept of monitoring window and expectation tolerance in Section \ref{sec:modelling}.
Additionally, the authors take a continuous time approach contrasted with ours that uses discrete time advancements.
Similarly, they model the property of availability using a two-state model.

>>>>

\section{Related Work}
\label{ch04:sec:relatedwork}

@@

Hogben and Pannetrat examine in \cite{hogben2013defavail} the challenges of defining and measuring availability to support real-world service comparison and dispute resolution through SLAs.
They show how two examples of real-world SLAs would lead one service provider to report 0\% availability while another would report 100\% for the same system state history but using a different period of time.
The transparency that the authors attempt to reach is addressed in our work by the concept of monitoring window and expectation tolerance in Section \ref{ch04:sec:modelling}.
Additionally, the authors take a continuous time approach contrasted with ours that uses discrete time advancements.
Similarly, they model the property of availability using a two-state model.

### 2016-05-04 p03-ch01-2015-monitoring/sec045-defs.tex
Add publication information to chapter abstracts
Remove all `vspace`
Replace `wrapfigure` with `figure`
Fix ref/label for Chapter 04

\section{Distributed Monitoring Model}
\label{sec:modelling}

We introduce a distributed monitoring platform and its components and
discuss some underlying assumptions and definitions.
Further, we define the notion of service availability and service budget compliance.
%\begin{wrapfigure}{r}{0.3\textwidth}
%\vspace{-20pt}
%\begin{center}
%\includegraphics[scale=0.7]{figs/deployment}
%\end{center}
%\vspace{-20pt}
%\caption{{\footnotesize High-level deployment architecture}}
%\label{fig:deployment}
%\vspace{-10pt}
%\end{wrapfigure}
%Figure \ref{fig:deployment} presents a high-level overview of services deployed in a distributed environment (e.g. ``the cloud'').
In the deployment environment (e.g., ``the cloud''), every  server from the IaaS provider is  used for a \emph{single} service of a customer, such as
the Query Service API for a customer of SDL Fredhopper (c.f. Section \ref{sec:fredhopper_example}).
Typically, multiple servers are allocated to a single customer.
The number of servers allocated for a customer is not visible to the customer.
The customer uses a single endpoint - in the load balancer layer - to access all their services.

@@

We then develop a distributed monitoring platform that aims to optimize these service
characteristics in a deployment environment.
 % given in Figure \ref{fig:deployment}.
The monitoring platform works in two cyclic phases: \emph{observation} and \emph{reaction}.
The observation phase takes measurements on services in the deployment environment.
Subsequently, the corresponding levels of the service characteristics are calculated.
In the reaction phase, if needed, a platform API is utilized to make the necessary changes to the deployment environment (e.g. adjust the number of allocated resources) to optimize the service characteristics.
% In the monitoring platform, not all components reside in the same resource in the deployment environment; all communications and interactions are performed in an \emph{asynchronous} way.
The monitoring platform builds on top of a real-time extension of the 
actor-based language ABS~\cite{johnsen2012abs}.
To ensure non-intrusiveness of the monitor with the running service,
each monitor is an active object (actor) running on a separate resource
from that which runs the service itself, and the components of the monitoring platform communicate through \emph{asynchronous messages} with deadlines~\cite{johnsen2012modeling}.
% The following elements are the major high-level components of the monitoring platform:
%for a distributed deployment environment:

@@

%\begin{assm}[Unlimited Capacity]
%\label{def:platform:cap}
%We assume that the external infrastructure provider is capable to provision an \emph{unlimited} number of resources.
%\end{assm}

@@

%\begin{assm}[Single Resource Type]
%\label{def:single:resource:type}
%To simplify reasoning, we assume that all resources are of the \emph{same type}; i.e. they have the same computing power, memory, and IO capacity.
%For example, if we are using AWS,
%we could choose the instance type \jtt{m1.large} for all the services.
%\end{assm}

@@

%\begin{assm}[Resource Initialization Time $t_i$]
%\label{def:resource:init:time}
%We assume that every resource $r$ that is initialized is ready for use in at most $t_i$ amount of time.
%% The time $t_i$ is bounded by a finite constant value.
%\end{assm}

%Initialization time can vary between different resource types. The above Assumption \ref{def:single:resource:type} simplifies this: we can just choose the initialization time (a fixed constant) of the single resource type that was chosen.
%The initialization time is also part of the contract with the IaaS provider.
%As discussed above in Assumption \ref{def:single:resource:type}, when a resource is launched and initialized, it might take a different amount of time to be in a state that is \emph{operational} and ready to be used.
%Thus,
%We simplify the property of initialization time of a resource to be a fixed constant for the single resource type that is used in the context of this research.

@@

We denote by $r$ a resource which provides computational power and storage
and by $s$ a general abstraction of a service in the deployment environment. 
A service exposes an API that is accessible through a delivery layer, such as HTTP.
In our example, a service is the Query API (c.f. Section~\ref{sec:fredhopper_example}) that is accessible through a single HTTP endpoint.

In our framework, 
\emph{monitoring platform}  $P$ is responsible for (de-)allocation of resources for computation or storage.
We abstract from a specific implementation of the monitoring platform $P$ through an API in Listing~\ref{lst:platform}. 
% 
\lstset{language=java,aboveskip=-20pt,belowskip=-20pt}
\begin{wrapfigure}{r}{0.5\textwidth}
\vspace{-15pt}
\begin{center}
\begin{lstlisting}[mathescape,caption=Platform API,label=lst:platform]
interface Platform {
  void    allocate(Service s);
  void    deallocate(Service s);
  Number  getState(Service s);
  boolean verify$_\alpha$(Service s);
  boolean verify$_\beta$(Service s);
}
\end{lstlisting}
\end{center}   
\vspace{-15pt}
\end{wrapfigure}
% 
There is only \emph{one} instance of $P$ available.
In this paper, $P$ internally uses an external infrastructure provisioning API to provide resources (e.g. AWS EC2).
The term ``platform'' is interchangeably used for monitoring in this paper.
The platform provides a method \jtt{getState(Service s)} which returns
the number of resources allocated to the given service $s$ at time $t_c$.

@@

\begin{defn}[Service Availability $\alpha(s,\tau,t_c)$]
\label{def:service:availability}
% 
First, we need a few auxiliary definitions before we can define service availability.
%

@@

We define the availability of a service $\alpha(s,\tau,t_c)$ in every monitoring window $\tau$ as:
\vspace{-10pt}
\[
\alpha(s,\tau,t_c) = \frac{\kappa_\sigma(s,\tau)}{E(s,\tau,t_c)}
\]
% in every monitoring window $\tau$.
%

@@

\begin{defn}[Budget Compliance $\beta(s,\tau)$]
\label{def:budget:compliance}
% 
We first provide a few auxiliary definitions.
%

>>>>

\section{Distributed Monitoring Model}
\label{ch04:sec:modelling}

We introduce a distributed monitoring platform and its components and
discuss some underlying assumptions and definitions.
Further, we define the notion of service availability and service budget compliance.
%\begin{wrapfigure}{r}{0.3\textwidth}
%\vspace{-20pt}
%\begin{center}
%\includegraphics[scale=0.7]{figs/deployment}
%\end{center}
%\vspace{-20pt}
%\caption{{\footnotesize High-level deployment architecture}}
%\label{ch04:fig:deployment}
%\vspace{-10pt}
%\end{wrapfigure}
%Figure \ref{ch04:fig:deployment} presents a high-level overview of services deployed in a distributed environment (e.g. ``the cloud'').
In the deployment environment (e.g., ``the cloud''), every  server from the IaaS provider is  used for a \emph{single} service of a customer, such as
the Query Service API for a customer of SDL Fredhopper (c.f. Section \ref{ch04:sec:fredhopper_example}).
Typically, multiple servers are allocated to a single customer.
The number of servers allocated for a customer is not visible to the customer.
The customer uses a single endpoint - in the load balancer layer - to access all their services.

@@

We then develop a distributed monitoring platform that aims to optimize these service
characteristics in a deployment environment.
 % given in Figure \ref{ch04:fig:deployment}.
The monitoring platform works in two cyclic phases: \emph{observation} and \emph{reaction}.
The observation phase takes measurements on services in the deployment environment.
Subsequently, the corresponding levels of the service characteristics are calculated.
In the reaction phase, if needed, a platform API is utilized to make the necessary changes to the deployment environment (e.g. adjust the number of allocated resources) to optimize the service characteristics.
% In the monitoring platform, not all components reside in the same resource in the deployment environment; all communications and interactions are performed in an \emph{asynchronous} way.
The monitoring platform builds on top of a real-time extension of the 
actor-based language ABS~\cite{johnsen2012abs}.
To ensure non-intrusiveness of the monitor with the running service,
each monitor is an active object (actor) running on a separate resource
from that which runs the service itself, and the components of the monitoring platform communicate through \emph{asynchronous messages} with deadlines~\cite{johnsen2012modeling}.
% The following elements are the major high-level components of the monitoring platform:
%for a distributed deployment environment:

@@

%\begin{assm}[Unlimited Capacity]
%\label{ch04:def:platform:cap}
%We assume that the external infrastructure provider is capable to provision an \emph{unlimited} number of resources.
%\end{assm}

@@

%\begin{assm}[Single Resource Type]
%\label{ch04:def:single:resource:type}
%To simplify reasoning, we assume that all resources are of the \emph{same type}; i.e. they have the same computing power, memory, and IO capacity.
%For example, if we are using AWS,
%we could choose the instance type \jtt{m1.large} for all the services.
%\end{assm}

@@

%\begin{assm}[Resource Initialization Time $t_i$]
%\label{ch04:def:resource:init:time}
%We assume that every resource $r$ that is initialized is ready for use in at most $t_i$ amount of time.
%% The time $t_i$ is bounded by a finite constant value.
%\end{assm}

%Initialization time can vary between different resource types. The above Assumption \ref{ch04:def:single:resource:type} simplifies this: we can just choose the initialization time (a fixed constant) of the single resource type that was chosen.
%The initialization time is also part of the contract with the IaaS provider.
%As discussed above in Assumption \ref{ch04:def:single:resource:type}, when a resource is launched and initialized, it might take a different amount of time to be in a state that is \emph{operational} and ready to be used.
%Thus,
%We simplify the property of initialization time of a resource to be a fixed constant for the single resource type that is used in the context of this research.

@@

We denote by $r$ a resource which provides computational power and storage
and by $s$ a general abstraction of a service in the deployment environment. 
A service exposes an API that is accessible through a delivery layer, such as HTTP.
In our example, a service is the Query API (c.f. Section~\ref{ch04:sec:fredhopper_example}) that is accessible through a single HTTP endpoint.

In our framework, 
\emph{monitoring platform}  $P$ is responsible for (de-)allocation of resources for computation or storage.
We abstract from a specific implementation of the monitoring platform $P$ through an API in Listing~\ref{ch04:lst:platform}. 
%

\lstset{language=java,aboveskip=-20pt,belowskip=-20pt}
\begin{figure}[h]
% \begin{wrapfigure}{r}{0.5\textwidth}
% \vspace{-15pt}
\begin{center}
\begin{lstlisting}[mathescape,caption=Platform API,label=lst:platform]
interface Platform {
  void    allocate(Service s);
  void    deallocate(Service s);
  Number  getState(Service s);
  boolean verify$_\alpha$(Service s);
  boolean verify$_\beta$(Service s);
}
\end{lstlisting}
\end{center}   
% \vspace{-15pt}
% \end{wrapfigure}
\end{figure}
\lstset{mathescape=false}
%

There is only \emph{one} instance of $P$ available.
In this paper, $P$ internally uses an external infrastructure provisioning API to provide resources (e.g. AWS EC2).
The term ``platform'' is interchangeably used for monitoring in this paper.
The platform provides a method \jtt{getState(Service s)} which returns
the number of resources allocated to the given service $s$ at time $t_c$.

@@

\begin{defn}[Service Availability $\alpha(s,\tau,t_c)$]
\label{ch04:def:service:availability}
% 
First, we need a few auxiliary definitions before we can define service availability.
%

@@

We define the availability of a service $\alpha(s,\tau,t_c)$ in every monitoring window $\tau$ as:
% \vspace{-10pt}
\[
\alpha(s,\tau,t_c) = \frac{\kappa_\sigma(s,\tau)}{E(s,\tau,t_c)}
\]
% in every monitoring window $\tau$.
%

@@

\begin{defn}[Budget Compliance $\beta(s,\tau)$]
\label{ch04:def:budget:compliance}
% 
We first provide a few auxiliary definitions.
%
